---
title: "BellaBeat - Data Analysis"
author: "Aidan Wilson"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  markdown: 
    wrap: 76
---

# BellaBeat - Data Analysis Capstone project

## By Aidan Wilson

Ask Prepare Process Analyze Share Act

### Characters and products

-   Urška Sršen: Bellabeat's cofounder and Chief Creative Officer
-   Sando Mur: Mathematician and Bellabeat's cofounder; key member of the
    Bellabeat executive team
-   Bellabeat marketing analytics team: A team of data analysts responsible
    for collecting, analyzing, and reporting data that helps guide
    Bellabeat's marketing strategy. You joined this team six months ago and
    have been busy learning about Bellabeat''s mission and business goals
    --- as well as how you, as a junior data analyst, can help Bellabeat
    achieve them.

### Product

-   Bellabeat app: The Bellabeat app provides users with health data related
    to their

    -   activity,
    -   sleep,
    -   stress,
    -   menstrual cycle,
    -   mindfulness habits. This data can help users better understand their
        current habits and make healthy decisions. The Bellabeat app
        connects to their line of smart wellness products.

-   Leaf: Bellabeat's classic wellness tracker can be worn as a bracelet,
    necklace, or clip. The Leaf tracker connects to the Bellabeat app to
    track activity, sleep, and stress.

-   Time: This wellness watch combines the timeless look of a classic
    timepiece with smart technology to track user:

    -   activity
    -   sleep
    -   stress The Time watch connects to the Bellabeat app to provide you
        with insights into your daily wellness.

-   Spring: This is a water bottle that tracks daily water intake using
    smart technology to ensure that you are appropriately hydrated
    throughout the day. The Spring bottle connects to the Bellabeat app to
    track your hydration levels.

-   Bellabeat membership: Bellabeat also offers a subscription-based
    membership program for users. Membership gives users 24/7 access to
    fully personalized guidance on nutrition, activity, sleep, health and
    beauty, and mindfulness based on their lifestyle and goals.

# Ask

Sršen asks you to analyze smart device usage data in order to gain insight
into how consumers use non-Bellabeat smart devices. She then wants you to
select one Bellabeat product to apply these insights to in your
presentation. These questions will guide your analysis:

1.  What are some trends in smart device usage?
2.  How could these trends apply to Bellabeat customers?
3.  How could these trends help influence Bellabeat marketing strategy?

# Deliverables

1.  A clear summary of the business task
2.  A description of all data sources used
3.  Documentation of any cleaning or manipulation of data
4.  A summary of your analysis
5.  Supporting visualizations and key findings 6.Your top high-level content
    recommendations based on your analysis

# Ask Challange:

Clear Statement of the business task: *"BellaBeat would like to see insights
into their smart devices, with this capstone focusing on one device. The
data and graphics should help convey their clientels current purchasing and
living habits and recommend BellaBeats next steps in marketing."*

Key stakeholders:

-   Urška Sršen: Co-founder and CCO
-   Sando Mur: BellaBeat exuctive team member
-   Marketing team

----------------------------------------------------------------------------

# Prepare

Look into the fitbit dataset

```{r}
library(tidyverse)
library(data.table)
library(readr)
library(dplyr)
library(broom)
library(psych) 
library(gt)
library(glue)
library(dplyr)
library(magrittr)
library(gtExtras)
library(lubridate)
library(RColorBrewer)

nb.cols <- 24
mycolors <- colorRampPalette(brewer.pal(8, "Paired"))(nb.cols)

file_location <- "/Users/aidan/Desktop/Capstone Project - Data Analysis/Fitabase Data 4.12.16-5.12.16/"
setwd("/Users/aidan/Desktop/Capstone Project - Data Analysis/Fitabase Data 4.12.16-5.12.16/")
#knitr::opts_knit$set(root.dir = "/Users/aidan/Desktop/Capstone Project - Data Analysis/Fitabase Data 4.12.16-5.12.16/")
Fit_Bit_File_Names <- list.files("/Users/aidan/Desktop/Capstone Project - Data Analysis/Fitabase Data 4.12.16-5.12.16/")

BellaBeat_Data <- lapply(paste0(file_location,Fit_Bit_File_Names),function(i){
  read.csv(i, header=TRUE)
})

#Looking at interesting file names to decide which dataset to work through
Fit_Bit_File_Names 

#Get an idea of what the data set looks like
head(BellaBeat_Data[[17]]) %>%
  gt()

#Get an idea of the sizes
dim(BellaBeat_Data[[17]])

```

Now that I have looked into the Data sets provided, based on what datasets
exist. I will choose to use the sleeping patterns data set.

I chose this as it posed the following issues:

-   Dealing with time and dates

-   Showing issues with lack of units being present

Now to clean up the data

```{r}

Sleep_Data <- BellaBeat_Data[[17]] %>%
  separate(SleepDay, sep = " ", into = c("Date", "Time", "AM or PM"))%>%
  mutate(TotalHoursAsleep = round(TotalMinutesAsleep/60, digits = 2))

#Now lets mess around with the table to get some information in usable formates. 

# 1) How many hours does each user sleep on average
Avg_rest_per_User <- Sleep_Data %>%
  group_by(Id) %>%
  summarise(mean(TotalHoursAsleep))%>%
  mutate(Avg_Sleep = round(as.numeric(`mean(TotalHoursAsleep)`),2)) %>%
  arrange(desc(Avg_Sleep))

# 2) Difference between the Hours alseep and Hours in bed
Bed_VS_Sleep <- Sleep_Data %>%
  mutate(Bed_VS_Sleep = round((TotalTimeInBed - TotalMinutesAsleep)/60, digits = 2))

```

Now that we have the data in a better orientation and legible formats we can
plot some graphs. hhhh

```{r}
ggplot(Avg_rest_per_User) +
  geom_bar(aes(x= reorder(as.character(Id), -Avg_Sleep),  y = Avg_Sleep, 
               fill = factor(as.character(Id))), stat = "identity", show.legend = FALSE)+
  scale_fill_manual(values = mycolors) +
  theme(
    plot.title = element_text("BellaBeat - Average Sleep per Person", size = 16, face = "bold"),
    axis.title.x = element_text("Individuals - By ID", size = 12, face = "italic"),
    axis.title.y = element_text("Average Sleep (Hrs)", size = 12, face = "italic"),
    axis.text.x = element_text(angle = 45, hjust=1)
    )
```
